### 2015-11-04 paper.Rmd
plots...

```{r minke-data-head}
library(Distance2)
head(minke)
```
and the amakihi data:
```{r amakihi-data-head}
head(amakihi)
```

@@

@@

@@

## Detection functions: key function plus adjustments (K+A)

@@

## half-normal
g.hn <- function(x,sigma) exp(-x^2/(2*sigma^2))
x <- seq(0,1,len=200)
for(this.sig in c(0.05, 0.25, 1, 10)){
  plot(x, g.hn(x,this.sig), main=bquote(sigma == .(this.sig)),
       xlab="", ylab="",
       type="l", xlim=c(0,1), ylim=c(0,1), asp=1)
  title(xlab="Distance", line=2)
  title(ylab="Detection probability", line=2)
}

## hazard-rate
g.hr <- function(x, sigma, b) 1 - exp(-(x/sigma)^-b)
for(this.sig in c(0.1, 0.5)){
  for(this.b in c(5, 1)){
    plot(x, g.hr(x, this.sig, this.b),
         main=bquote(sigma == .(this.sig) ~ .(", b") == .(this.b)),
         xlab="", ylab="",
         type="l", xlim=c(0,1), ylim=c(0,1), asp=1)
    title(xlab="Distance", line=2)
    title(ylab="Detection probability", line=2)
  }
}

@@

## Detection functions: mixture models

@@

## Covariates

@@

@@

* writing new detection functions
* Interface to dsm etc

>>>>

```{r minke-data-head, message=FALSE}
library(Distance2)
head(minke)
```
and the amakihi data:
```{r amakihi-data-head}
head(amakihi)
```

@@

It is important that the detection function accurately models detectability near zero distance. We are less worried by its behaviour further away from 0. To ensure that the model is not overly influence by distances far from zero we truncate the distances beyond a given distance $w$ (known as the *truncation distance*). Further details on truncation can be found in **CITATION**.

@@

Given these criteria, we can start to think about models for $g$.

## Formulations

@@

### Key function plus adjustments (K+A)

@@

## half-normal
g.hn <- function(x,sigma) exp(-x^2/(2*sigma^2))
for(this.sig in c(0.05, 0.25, 1, 10)){
  this.g <- function(x) g.hn(x, sigma=this.sig)
  curve(this.g, from=0, to=1, xlab="", ylab="",
        main=bquote(sigma == .(this.sig)),
        xlim=c(0,1), ylim=c(0,1), asp=1)
  title(xlab="Distance", line=2)
  title(ylab="Detection probability", line=2)
}

## hazard-rate
g.hr <- function(x, sigma, b) 1 - exp(-(x/sigma)^-b)
for(this.sig in c(0.1, 0.5)){
  for(this.b in c(5, 1)){
    this.g <- function(x) g.hr(x, sigma=this.sig, b=this.b)
    curve(this.g, from=0, to=1, xlab="", ylab="",
          main=bquote(sigma == .(this.sig) ~ .(", b") == .(this.b)),
          xlim=c(0,1), ylim=c(0,1), asp=1)
    title(xlab="Distance", line=2)
    title(ylab="Detection probability", line=2)
  }
}

@@

```{r adjust-mix-comp, fig.width=10, fig.height=5, fig.cap="Top row shows possible shapes for the detection function when adjustments are included for half-normal and hazard-rate models. Bottom row shows possible shapes when 2-component mixtures of half-normal functions are used. Notation is as in main text.", echo=FALSE}
par(mfrow=c(2,4), mar=c(3.5, 3, 2, 1) + 0.1)

## half-normal with cosine
g <- function(x) exp(-x^2/(2*0.01^2))*(1+0.5*cos((2*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Half-normal with 1 cosine adjustment",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)
# 2 cosines
g <- function(x) exp(-x^2/(2*0.01^2))*(1-0.06*cos((2*pi*x)/0.025)+0.25*cos((3*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Half-normal with 2 cosine adjustments",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)

## hazard-rate with cosine
g <- function(x) (1 - exp(-(x/0.005)^-1.1))*(1-0.05*cos((2*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Hazard-rate with 1 cosine adjustment",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)
# 2 cosines
g <- function(x) (1 - exp(-(x/0.005)^-1.9))*
                  (1+0.05*cos((2*pi*x)/0.025) + 0.25*cos((3*pi*x)/0.025))
f <- function(x) g(x)/g(0)
curve(f, from=0, to=0.025, xlab="", ylab="", main="Hazard-rate with 2 cosine adjustments",xlim=c(0,0.025), ylim=c(0,1))
title(xlab="Distance", line=2)
title(ylab="Detection probability", line=2)

# mixture models
plotmix <- function(sig1, sig2, mix){

  gmix <- function(x) mix*g.hn(x, sig1) + (1-mix)*g.hn(x,sig2)

  curve(gmix, from=0, to=1, xlab="", ylab="",
        main=bquote(sigma[1] == .(sig1) ~ sigma[2] == .(sig2) ~  phi[1] == .(mix)),
        xlim=c(0,1), ylim=c(0,1), asp=1)
  title(xlab="Distance", line=2)
  title(ylab="Detection probability", line=2)
}
plotmix(0.8, 0.15, 0.3)
plotmix(0.6, 0.1,0.7)
plotmix(10, 0.2, 0.15)
plotmix(0.7, 0.05, 0.6)

### Mixture models

@@

### Covariates

@@

## Fitting detection functions in R

The workhorse of detection function fitting in \pkg{Distance2} is the \code{ds} function. Here we show off the formulations for the detection function that we've seen above for both the minke whale and amakihi data.

We can fit a model to the minke whale data, setting the truncation at 1.5km and using the default options in \code{ds} very simply:
```{r minke-hn}
minke_hn <- ds(minke, truncation=1.5)
```
**Figure???** shows the result of calling \code{plot} on the resulting model object.

```{r minke-hn-plot, fig.width=4, fig.height=4, fig.cap="Plot of fitted detection function overlayed on the histogram of observed distances for the minke whale data."}
plot(minke_hn)
```

@@

* writing new detection functions
* Interface to dsm etc
* MRDS?
* why is + important?
